Two Mode Networks and Relational Similarity

### 2025-01-27 Reviews and Response/response-memo.tex

\textbf{Response}:

\underline{Data errors in the Southern Women two-mode matrix}:

\textbf{Reviewer}: \textit{Importantly, there are 4 errors in the matrix (``Southern Women data") that you analyze. For this dataset, you (in the note to Table 1) cite Doreian et al (2004, Table 4). Among the attendees at Event E4, you show Frances, Eleanor, and Ruth. However they did not attend this event. You also show Nora as absent from Event E8; however, Nora attended E8. These four errors are discrepant in comparison to your cited source (Doreian et al, 2004, Table 4), and they are discrepant as well in comparison to the original source (Davis, Gardener, and Gardener, 1941, p. 141, which has events in the same order as in your Table 1, and a slightly different order of women; this table is reproduced in Freeman, 2003, p. 42).}

\begin{figure}[h!]
    \includegraphics[width=1.0\textwidth]{Reviews and Response/sw-original.png}
\end{figure}

\textbf{Response}: Thanks for calling attention to this---rather embarrassing---error; it's definitely a cautionary tale against manually copying data. I have corrected the data, and as you can see in the analysis file included as supplementary material, I now extract the Southern Women data directly from the \texttt{networkdata} package. My table now matches those in the references you mentioned exactly. Note, however, that while you are correct that Frances, Eleanor, and Ruth did not attend E4, Nora did not attend E8, at least according to the published tables in Doreian et al (2004), Kovacs (2010), and the Figure 1 from Freeman (2003), shown above; in these papers, as you can see, Nora's event set is \{E6, E7, E9, E10, E11, E12, E13, E14\}. I think the confusion comes from the fact that Freeman used a non-chronological order of events (as is evident from the date labels in the figure) that emphasizes the group structure of the women. This non-chronological ordering leads to a different labeling of events than when events are arranged in chronological order; this is the ordering of events used by Doreian et al. 2004 and Kovacs (2010). In this event labeling, Nora does not attend E8 (the event that took place on 9/16.) However, note that this is obviously \textit{not} the ordering of events in the original Breiger (1974)---nor for that matter the earlier Doreian (1979) piece in \textit{Social Networks} that used Q-Analysis. In this \textit{chronological} labeling of events, Nora does attend ``E8" (the event that took place on 5/19), as evident in Breiger (1974, p. 186, fig. 2a). In any case, to forestall any confusion, and given that the in the \texttt{networkdata} package, the \texttt{igraph} object containing the Southern Women data come ordered using the Freeman (2003) ordering, I'm sticking to the non-chronological event labeling. I include a very brief discussion in the paper clarifying this issue.

### 2025-01-27 main-R1.tex
changed paper

\subsection{Empirical Example} \label{subsec:ex}
Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons in the Southern Women data \citep{davis1941}, using a bump chart to plot the rank order trajectory of persons across reflections. The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis. As the figure shows, $NORA$, $FLORA$, $CHARLOTTE$, and $EVELYN$ are the top-ranked actors when it comes to $C^R_p(2)$:  The average number of members of the groups they belong to. However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections). Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover the Doreian et al. (2004) block partition (see Table~\ref{tab:southern}), but this time with Dorothy and Pearl in the middle separating the two largest blocks of women.

>>>>

\subsection{Empirical Example} \label{subsec:ex}
The following analyses use the classic Southern Women data \citep{davis1941} as the running example, a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten longer since his analytic ``meta-analysis'' \citep[e.g.,][]{doreian2004generalized, field2006identifying, wang2009exponential, everett2018measuring, batagelj2022analysis, lizardo2024two}. As we can see, novel insights can still be garnered from these old data.

Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons, using a bump chart to plot the rank order trajectory of persons across reflections. The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis. As the figure shows, $NORA$, $FLORA$, $CHARLOTTE$, and $EVELYN$ are the top-ranked actors when it comes to $C^R_p(2)$:  The average number of members of the groups they belong to. However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections). Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover the Doreian et al. (2004) block partition (see Table~\ref{tab:southern}), but this time with Dorothy and Pearl in the middle separating the two largest blocks of women.
